---
title: "Estimates of pi0"
author: "David Gerard"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
fig_caption: yes
---

# Abstract
I plot estimates of $\pi_0$ versus $\pi_0$ for various methods under various alternative densities. RUVASH and SUCCOTASH perform the best.

# Method
I used $\pi_0$ from a grid from $0.01$ to $0.99$, generated count data from the muscle GTEX data with $p = 1000$, where $p$ is the number of genes. Signal was added via the Poisson thinning procedure with log2-fold-changes drawn from the same alternative distributions as in the ASH paper (see table below). The only difference is that the mixing variances were also divided by $n - 2$, where $n$ is the sample size. This is a little different from my previous simulation settings and is meant to conform to the settings that Mengyin has been looking at.

\begin{center}
\begin{tabular}{ll}
Scenario & Alternative Distribution \\
\hline
Spiky & $0.4N(0, 0.25^2) + 0.2N(0, 0.5^2) + 0.2N(0, 1^2), 0.2N(0, 2^2)$\\
Near Normal & $2/3N(0, 1^2) + 1/3N(0, 2^2)$\\
Flattop & $(1/7)N(-1.5, .5^2) + N(-1, .5^2) + N(-.5, .5^2)+N(0, .5^2) + N(0.5, .5^2) + N(1.0, .5^2) + N(1.5, .5^2)$\\
Skew & $(1/4)N(-2, 2^2) + (1/4)N(-1, 1.5^2) + (1/3)N(0, 1^2) + (1/6)N(1, 1^2)$\\
Big-normal & $N(0, 4^2)$\\
Bimodal & $0.5N(-2, 1^2) + 0.5N(2, 1^2)$
\end{tabular}
\end{center}

I ran 10 methods, keeping their estimates of $\pi_0$:

* OLS + qvalue.
* OLS + ASH
* Two-step SUCCOTASH using normal mixtures and heteroscedastic PCA as the factor-analysis method.
* The robust regression version of CATE using PCA as the factor analysis method + qvalue.
* SVA + qvalue.
* Scaled version of RUVASH.
* Negative control version of CATE using PCA as the factor analysis method + qvalue.
* RUV2 + qvalue.
* RUV4 + qvalue.
* Ridge version of LEAPP + qvalue.

Plots of $\pi_0$ vs $\hat{\pi}_0$ for each method in each combination of $n$ by alternative type are plotted below.

# Results
SUCCOTASH and RUVASH seem to perform the best, but all of the methods seem to be getting better at estimating $\pi_0$ as $n$ increases.

In this case where the mixing variances decrease with $n$, SUCCOTASH and RUVASH actually perform the worst in the spiky scenario rather than the bimodal scenario. But they perform better than the other methods in the spiky scenario.

```{r, echo = FALSE, warning = FALSE, cache = TRUE, fig.height = 9}
library(ggplot2)
library(reshape2)
library(ggsci)
pi0hat <- read.csv("pi0hat_df.csv")
mypal = pal_npg()(10)
for (index in 1:10) {
    p <- ggplot(data = pi0hat, mapping = aes(x = nullpi, y = pi0hat[, index],
                                             color = I(mypal[index]))) +
        facet_grid(alt_type~Nsamp) +
        geom_point(pch = I(16)) +
        geom_abline(slope = 1, intercept = 0) +
        ylab(expression(hat(pi)[0])) +
        xlab(expression(pi[0])) +
        ggtitle(paste(names(pi0hat)[index], "pi0hat.")) +
        theme_bw() +
        theme(strip.background = element_rect(fill = "white"))
    print(p)
}
```

```{r}
sessionInfo()
```
