---
title: "Different Alternative Types"
author: "David Gerard"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
bibliography: sva_bib.bib
fig_caption: yes
---



# Abstract
I compare SUCCOTASH to various competitors under the same alternative scenarios in @stephens2016false. SUCCOTASH generally has superior performance in terms of (1) estimating $\pi_0$, (2) having higher AUC, and (3) having lower MSE. LEAPP does better under some scenarios in terms of estimating $\pi_0$ but has far worse AUC and MSE. SUCCOTASH accurately estimates the average sign error rate, though not in as conservative manner as one might hope in some scenarios.


# Simulation Setup
I ran through 200 repetitions of generating data from GTEX muscle data under the following parameter conditions:

* $n \in \{10, 20, 40\}$,
* $p = 1000$.
* $\pi_0 \in \{0.5, 0.9\}$,
* The alternative distribution being either spiky, near-normal, flattop, skew, big-normal, or bimodal, where these are the same alternatives defined in @stephens2016false and the following table. New alternatives are generated every iteration.

\begin{center}
\begin{tabular}{ll}
Scenario & Alternative Distribution \\
\hline
Spiky & $0.4N(0, 0.25^2) + 0.2N(0, 0.5^2) + 0.2N(0, 1^2), 0.2N(0, 2^2)$\\
Near Normal & $2/3N(0, 1^2) + 1/3N(0, 2^2)$\\
Flattop & $(1/7)N(-1.5, .5^2) + N(-1, .5^2) + N(-.5, .5^2)+N(0, .5^2) + N(0.5, .5^2) + N(1.0, .5^2) + N(1.5, .5^2)$\\
Skew & $(1/4)N(-2, 2^2) + (1/4)N(-1, 1.5^2) + (1/3)N(0, 1^2) + (1/6)N(1, 1^2)$\\
Big-normal & $N(0, 4^2)$\\
Bimodal & $0.5N(-2, 1^2) + 0.5N(2, 1^2)$
\end{tabular}
\end{center}

I extracted the most expressed $p$ genes from the GTEX muscle data and $n$ samples are chosen at
random. Half of these samples are randomly given the "treatment"
label 1, the other half given the "control" label 0. Of the $p$
genes, $\pi_0p$ were chosen to be non-null. Signal was added by a
Poisson-thinning approach, where the log-2 fold change was sampled from one of five the alternative models above.
That is
\begin{align}
  A_1,\ldots,A_{p/2} &\sim f\\
  B_i &= 2^{A_i} \text{ for } i = 1,\ldots, p/2,
\end{align}
where $f$ is from the table above. If $A_i > 0$ then we replace $Y_{[1:(n/2), i]}$ with $Binom(Y_{[j,i]}, 1 / B_i)$ for $j = 1,\ldots, n/2$. If $A_i < 0$ then we replace $Y_{[(n/2 + 1):n, i]}$ with $Binom(Y_{[j,i]}, B_i)$ for $j = n/2 + 1,\ldots, n$.

I now describe the justification for this. Suppose that
\begin{align}
Y_{ij} \sim Poisson(\lambda_j).
\end{align}
Let $x_{i}$ be the indicator of treatment vs control for individual $i$. Let $\Omega$ be the set of non-null genes. Let $Z$ be the new dataset derived via the steps above. That is
\begin{align}
Z_{ij} | Y_{ij}=
\begin{cases}
Binom(Y_{ij}, 2 ^ {A_{j}x_i}) & \text{if } A_{j} < 0 \text{ and } j \in \Omega\\
Binom(Y_{ij}, 2 ^ {-A_{j}(1 - x_i)}) & \text{if } A_{j} > 0 \text{ and } j \in \Omega\\
Y_{ij} & \text{if } j \not\in \Omega.
\end{cases}
\end{align}
Then
\begin{align}
Z_{ij} | A_{j}, A_{j} < 0, j \in \Omega &\sim Poisson(2^{A_{j}x_i}\lambda_j)\\
Z_{ij} | A_{j}, A_{j} > 0, j \in \Omega &\sim Poisson(2^{-A_{j}(1 - x_i)}\lambda_j),
\end{align}
and
\begin{align}
E[\log_2(Z_{ij}) - \log_2(Z_{kj}) | A_{j}, A_{j} < 0, j \in \Omega] &\approx A_{j}x_i - A_{j}x_k, \text{ and}\\
E[\log_2(Z_{ij}) - \log_2(Z_{kj}) | A_{j}, A_{j} > 0, j \in \Omega] &\approx -A_{j}(1 - x_i) + A_{j}(1 - x_k).
\end{align}
if individual $i$ is in the treatment group and individual $k$ is in the control group, then this just equals $A_j$. I treat the $A_j$'s as the true coefficient values when calculating the MSE below.

# Methods
I first normalized the counts by $\log_2(COUNTS + 1)$. The number of hidden confounders was estimated using the methods of @buja1992remarks implemented in the \texttt{num.sv()} function in the \texttt{sva} package in \texttt{R}.

The confounder adjustment methods I look at in this write-up are:

* OLS + qvalue.
* RUVols + estimate variance inflation using controls + ASH
* RUVgls + estimate variance inflation using controls + ASH
* RUVgls + estimate variance inflation using controls + MLE to UMVUE motivated scaling + ASH
* SUCCOTASH using normal mixtures and heteroscedastic PCA as the factor-analysis method.
* The robust regression version of CATE using PCA as the factor analysis method + qvalue.
* SVA + qvalue.
* Negative control version of CATE using PCA as the factor analysis method + qvalue.
* RUV2 + qvalue.
* RUV4 + qvalue.
* Sparse version of LEAPP. Since this is a sparsity-inducing procedure, I used the proportion of zeros as the estimate of $\pi_0$.
* Ridge version of LEAPP + qvalue.

# RUVASH
The model for the second step in RUV4, LEAPP, CATE, and SUCCOTASH is
\begin{align}
Y &= \beta_{p \times 1} + \hat{\alpha}_{p \times q}Z_{q \times 1} + E_{p \times 1}\\
E &\sim N_p(0, \lambda\hat{\Sigma})\\
\hat{\Sigma} &= diag(\hat{\sigma}_1^2,\ldots,\hat{\sigma}_p^2),
\end{align}
with $\hat{\Sigma}$ and $\hat{\alpha}$ known. $\lambda$ is a variance inflation parameter that is novel to SUCCOTASH.

Let $\mathcal{C} \subseteq \{1,\ldots,p\}$ denote the indices for the negative controls. I.e. $\beta_{\mathcal{C}} = 0$. Then the model for the negative controls is
\begin{align}
Y_{\mathcal{C}} = \alpha_{\mathcal{C}}Z + E_{\mathcal{C}}.
\end{align}
The maximum likelihood estimates of $Z$ and $\lambda$ are easy:
\begin{align}
\hat{Z} = (\alpha_{\mathcal{C}}^T\hat{\Sigma}^{-1}\alpha_{\mathcal{C}})^{-1}\alpha_{\mathcal{C}}^T\hat{\Sigma}^{-1}Y_{\mathcal{C}}.
\end{align}
Let $R_{\mathcal{C}} = Y_{\mathcal{C}} - \hat{\alpha}_{\mathcal{C}}\hat{Z}$. Then the MLE of $\lambda$ is
\begin{align}
\hat{\lambda} = \frac{1}{|\mathcal{C}|}\sum_{i \in \mathcal{C}} R_{i}^2/\hat{\sigma}_i^2.
\end{align}

RUVASH runs ASH with means $\hat{\beta} = Y - \hat{\alpha}\hat{Z}$ and variances $\hat{\lambda}\hat{\sigma}_i^2$.

In the same way that I made an ad-hoc correction to the SUCCOTASH variance inflation parameter, I also explore using this for RUVASH. Specifically, I set
\begin{align}
\tilde{\lambda} = \frac{n}{n - k - q} \hat{\lambda}
\end{align}
and run ASH with variances $\tilde{\lambda}\hat{\sigma}_i^2$. These are called SRUVASH in the plots below.



# Results
Note that in the plots below, $n$ refers to the size of each group, not the total size.

## Estimates of $\pi_0$

* RUV + ASH Works much better here than vanilla ASH at estimating $\pi_0$, but it is slightly anti-conservative.
* RUV + ad-hoc scaling + ASH has conservative FDR with high probability in almost every scenario. The only one where it isn't conservative is the Bimodal alternative scenario.
* SUCCOTASH has slightly anti-conservative estimates of $\pi_0$ in the Flattop and bimodal Scenarios. It does well for every other scenario for larger n.
* LEAPP does amazingly well in the Big-normal and bimodal scenarios, even for $n = 10$. However, it is far too conservative in every other non-null scenario. This seems to indicate that LEAPP functions best if there is a separation of the alternative signal from zero.
* No method using qvalue ever performed as well as succotash. Indeed, none exhibited this "conservative with high-probability" behavior that is desirable.


## AUC performance.

* SUCCOTASH always has higher AUC, even in the bimodal and big-normal scenarios where LEAPP estimated $\pi_0$ more accurately.
* All of the RUV + ASH methods have about the same AUC as SUCCOTASH --- including the one with the ad-hoc scaling value that works so well with estimating $\pi_0$.

## ROC
I calculated the point-wise mean true positive rate (TPR) and false positive rate (FPR) at each ordered position under each scenario. These are plotted after the AUC boxplots. SUCCOTASH and the RUV + ASH methods always have higher and further left point-wise average ROC curves, even under the bimodal scenario.

## MSE

* SUCCOTASH has superior performance in term of MSE compared to other methods.
* LEAPP has terrible MSE performance in all cases except the all-null setting.
* Even then, SUCCOTASH has better performance.
* RUV + ASH has pretty good MSE performance.
* The ad-hoc scaling version of RUV + ASH seems to suffer in MSE a little bit, especially at low sample sizes.


\clearpage


```{r, cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(pROC)



## pi0hat plots ------------------------------------------------------------------------
load("pi0hat_muscle.Rd")
load("method_vec.Rd")
method_vec <- c(method_vec, "RUVASHOLS", "RUVASHGLS", "SRUVASHGLS")
load("pi0hat_muscle_ash.Rd")
load("pi0hat_muscle_ruvash.Rd")
load("pi0hat_muscle_sruvash.Rd")


for (index in 1:length(ash_pi0hat)) {
    pi0hat[[index]] <- c(pi0hat[[index]], ash_pi0hat[[index]], ruvash_pi0hat[[index]],
                         sruvash_pi0hat[[index]])
}

par_vals <- read.csv("sim_settings.csv")

nullpi_seq   <- unique(par_vals$nullpi)
nsamp_seq    <- unique(par_vals$Nsamp)
alt_type_seq <- unique(par_vals$alt_type)


for (alt_type_index in 1:(length(alt_type_seq) - 1)) {
    which_alt <- par_vals$alt_type == alt_type_seq[alt_type_index]

    pi0mat <- as.data.frame(t(sapply(pi0hat[which_alt], FUN = c)))
    names(pi0mat) <- method_vec
    pi0mat$nullpi <- par_vals$nullpi[which_alt]
    pi0mat$Nsamp  <- par_vals$Nsamp[which_alt]

    ## pi0_plot
    dummy_dat <- expand.grid(nullpi_seq[-length(nullpi_seq)], nsamp_seq)
    colnames(dummy_dat) <- c("nullpi", "Nsamp")

    name_vec <- colnames(pi0mat)
    colnames(pi0mat) <- gsub("pi0hat_", "", x = name_vec)
    long_dat <- melt(pi0mat, id.vars = c("nullpi", "Nsamp"),
                     measure.vars = colnames(pi0mat)[1:(ncol(pi0mat) - 2)])
    p <- ggplot(data = long_dat, mapping = aes(x = variable, y = value, fill = variable))
    p <- p + facet_grid(nullpi~Nsamp) + ylab(expression(hat(pi)[0]))
    p <- p + geom_boxplot(size = 0.4, outlier.size = 0.4)
    p <- p + geom_hline(data = dummy_dat, aes(yintercept = nullpi), lty = 2, color = "red", lwd = 1)
    p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))
    p <- p + ggtitle(paste("Estimates of pi0 When Using Muscle Tissue, Alternative =",
                           alt_type_seq[alt_type_index]))
    print(p)
}

which_alt <- par_vals$alt_type == "all_null"
pi0mat <- as.data.frame(t(sapply(pi0hat[which_alt], FUN = c)))
names(pi0mat) <- method_vec
pi0mat$nullpi <- par_vals$nullpi[which_alt]
pi0mat$Nsamp  <- par_vals$Nsamp[which_alt]
dummy_dat <- data.frame(Nsamp = nsamp_seq, nullpi = rep(1, length = length(nsamp_seq)))

name_vec <- colnames(pi0mat)
colnames(pi0mat) <- gsub("pi0hat_", "", x = name_vec)
long_dat <- melt(pi0mat, id.vars = c("Nsamp"),
                 measure.vars = colnames(pi0mat)[1:(ncol(pi0mat) - 2)])
p <- ggplot(data = long_dat, mapping = aes(x = variable, y = value, fill = variable))
p <- p + facet_grid(.~Nsamp) + ylab(expression(hat(pi)[0]))
p <- p + geom_boxplot(size = 0.4, outlier.size = 0.4)
p <- p + geom_hline(data = dummy_dat, aes(yintercept = nullpi), lty = 2, color = "red", lwd = 1)
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))
p <- p + ggtitle("Estimates of pi0 When Using Muscle Tissue and All Null")
print(p)
```

\clearpage

```{r, cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(pROC)

## AUC Boxplots ---------------------------------------------------------------------
rm(list = ls())
load("lfdr_muscle.Rd")
load("lfdr_muscle_ash.Rd")
load("lfdr_muscle_ruvash.Rd")
load("lfdr_muscle_sruvash.Rd")

for (index in 1:length(ash_lfdr)) {
    lfdr[[index]]$RUVASHOLS <- ash_lfdr[[index]][,1]
    lfdr[[index]]$RUVASHGLS <- ruvash_lfdr[[index]][,1]
    lfdr[[index]]$SRUVASHGLS <- sruvash_lfdr[[index]][,1]
    lfdr[[index]] <- lfdr[[index]][, c(1:(ncol(lfdr[[index]]) - 4),
                                       ncol(lfdr[[index]]) - 2, ncol(lfdr[[index]]) - 1,
                                       ncol(lfdr[[index]]),
                                       ncol(lfdr[[index]]) - 3)]
}

par_vals <- read.csv("sim_settings.csv")

get_auc <- function(lfdr_mat) {
    null_col <- which(colnames(lfdr_mat) == "which_null")
    auc_vec <- rep(NA, length = ncol(lfdr_mat) - 1)
    for(index in (1:length(auc_vec))[-null_col]) {
        if (all(is.na(lfdr_mat[, index]))) {
            auc_vec[index] <- NA
        } else {
            auc_vec[index] <- roc(response = lfdr_mat$which_null,
                                  predictor = c(lfdr_mat[, index]))$auc
            }
    }
    return(auc_vec)
}

## which_allnull <- par_vals$alt_type == "all_null"
## auc_mat <- as.data.frame(t(sapply(lfdr[!which_allnull], FUN = get_auc)))
## names(auc_mat) <- colnames(lfdr[[1]])[1:(ncol(lfdr[[1]]) - 1)]
## auc_mat$nullpi   <- par_vals$nullpi[!which_allnull]
## auc_mat$Nsamp    <- par_vals$Nsamp[!which_allnull]
## auc_mat$alt_type <- par_vals$alt_type[!which_allnull]
## write.csv(auc_mat, file = "auc_mat.csv", row.names = FALSE)

auc_mat <- read.csv("auc_mat.csv")
alt_type_seq <- unique(auc_mat$alt_type)


for (alt_type_index in 1:length(alt_type_seq)) {
    current_alt_type <- alt_type_seq[alt_type_index]

    auc_mat_sub <- auc_mat[auc_mat$alt_type == current_alt_type,]
    auc_mat_sub <- dplyr::select(auc_mat_sub, -alt_type)


    ## create dummy dat to plot max median auc
    nullpi_seq <- unique(auc_mat_sub$nullpi)
    nsamp_seq <- unique(auc_mat_sub$Nsamp)
    dummy_dat <- expand.grid(nullpi_seq, nsamp_seq)
    colnames(dummy_dat) <- c("nullpi", "Nsamp")
    med_mat <- matrix(NA, nrow = length(nullpi_seq) * length(nsamp_seq), ncol = ncol(auc_mat_sub) - 3)
    for (index in 1:(ncol(auc_mat_sub) - 3)) {
        form1 <- as.formula(paste(colnames(auc_mat_sub)[index], "~ nullpi + Nsamp"))
        out1 <- aggregate(form1, FUN = median, na.rm = TRUE,
                          data = auc_mat_sub)
        med_mat[, index] <- out1[, 3]
    }
    dummy_dat2 <- cbind(expand.grid(nullpi_seq[nullpi_seq != 1], nsamp_seq), apply(med_mat, 1, max))
    colnames(dummy_dat2) <- c("nullpi", "Nsamp", "max_med")
    ## end dummy dat


    long_dat <- melt(auc_mat_sub, id.vars = c("nullpi", "Nsamp"))
    p <- ggplot(data = long_dat, mapping = aes(x = variable, y = value, fill = variable))
    p <- p + facet_grid(nullpi~Nsamp) + ylab("AUC")
    p <- p + geom_boxplot(size = 0.4, outlier.size = 0.4)
    p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))
    p <- p + geom_hline(data = dummy_dat2, aes(yintercept = max_med), lty = 2,
                        color = "red", lwd = 0.6)
    p <- p + ggtitle(paste("AUC When Using Muscle Tissue, Alternative =",
                           alt_type_seq[alt_type_index]))
    print(p)
}
```

\clearpage

```{r, cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(pROC)
## Mean ROC Curves ------------------------------------------------------------------------
rm(list = ls())
load("lfdr_muscle.Rd")
load("lfdr_muscle_ash.Rd")
load("lfdr_muscle_ruvash.Rd")
load("lfdr_muscle_sruvash.Rd")
par_vals <- read.csv("sim_settings.csv")

for (index in 1:length(ash_lfdr)) {
    lfdr[[index]]$RUVASHOLS <- ash_lfdr[[index]][,1]
    lfdr[[index]]$RUVASHGLS <- ruvash_lfdr[[index]][,1]
    lfdr[[index]]$SRUVASHGLS <- sruvash_lfdr[[index]][,1]
    lfdr[[index]] <- lfdr[[index]][, c(1:(ncol(lfdr[[index]]) - 4),
                                       ncol(lfdr[[index]]) - 2, ncol(lfdr[[index]]) - 1,
                                       ncol(lfdr[[index]]),
                                       ncol(lfdr[[index]]) - 3)]
}

nullpi_seq <- unique(par_vals$nullpi)
Nsamp_seq <- unique(par_vals$Nsamp)
alt_type_seq <- unique(par_vals$alt_type)

get_fd <- function(predictor_vec, response_vec) {
    pord <- order(predictor_vec)
    oresp <- response_vec[pord]
    fd <- cumsum(oresp)
    return(fd)
}

get_roc <- function(lfdr_mat) {
    roc_out <- apply(lfdr_mat[-ncol(lfdr_mat)], 2, get_fd, response_vec = lfdr_mat$which_null)

    fpr <- roc_out / sum(lfdr_mat$which_null)
    tpr <- (1:nrow(roc_out) - roc_out) / sum(1 - lfdr_mat$which_null)

    ##matplot(fpr, tpr, type = "l", col = 1:9, lty = 1)
    ##legend("bottomright", colnames(lfdr_mat), col = 1:9, lty = 1)
    return(list(tpr = tpr, fpr = fpr))
}

roc_out <- lapply(lfdr, get_roc)


ft_prsum <- function(ftpr_list) {
    tpr_mean <- matrix(0, nrow = nrow(ftpr_list[[1]][[1]]), ncol = ncol(ftpr_list[[1]][[1]]))
    fpr_mean <- matrix(0, nrow = nrow(ftpr_list[[1]][[2]]), ncol = ncol(ftpr_list[[1]][[2]]))

    for (index in 1:length(ftpr_list)) {
        tpr_mean <- tpr_mean + ftpr_list[[index]][[1]]
        fpr_mean <- fpr_mean + ftpr_list[[index]][[2]]
    }
    tpr_mean <- tpr_mean / length(ftpr_list)
    fpr_mean <- fpr_mean / length(ftpr_list)
    return(list(tpr = tpr_mean, fpr = fpr_mean))
}


ave_ft_list <- list()
list_index <- 1
for (nullpi_index in 1:(length(nullpi_seq) - 1)) {
    nullpi_current <- nullpi_seq[nullpi_index]
    for(Nsamp_index in 1:length(Nsamp_seq)) {
        Nsamp_current <- Nsamp_seq[Nsamp_index]
        for(alt_type_index in 1:(length(alt_type_seq) - 1)) {
            alt_type_current <- alt_type_seq[alt_type_index]
            current_list <- par_vals$nullpi == nullpi_current & par_vals$Nsamp == Nsamp_current &
                par_vals$alt_type == alt_type_current
            ave_ft_list[[list_index]] <- ft_prsum(roc_out[current_list])
            list_index <- list_index + 1
        }
    }
}



summary_par_vals <- expand.grid(alt_type_seq[1:(length(alt_type_seq) - 1)], Nsamp_seq, nullpi_seq[1:(length(nullpi_seq) - 1)])
names(summary_par_vals) <- c("alt_type", "Nsamp", "nullpi")


list_index <- 1
fpr <- ave_ft_list[[list_index]][[2]]
tpr <- ave_ft_list[[list_index]][[1]]

temp_mat          <- cbind(melt(fpr)[, 2:3], melt(tpr)$value)
temp_mat$nullpi   <- summary_par_vals$nullpi[list_index]
temp_mat$Nsamp    <- summary_par_vals$Nsamp[list_index]
temp_mat$alt_type <- summary_par_vals$alt_type[list_index]
colnames(temp_mat) <- c("Method", "fpr", "tpr", "nullpi", "Nsamp", "alt_type")

roc_mat <- temp_mat
for (list_index in 2:nrow(summary_par_vals)) {
    fpr <- ave_ft_list[[list_index]][[2]]
    tpr <- ave_ft_list[[list_index]][[1]]

    temp_mat          <- cbind(melt(fpr)[, 2:3], melt(tpr)$value)
    temp_mat$nullpi   <- summary_par_vals$nullpi[list_index]
    temp_mat$Nsamp    <- summary_par_vals$Nsamp[list_index]
    temp_mat$alt_type <- summary_par_vals$alt_type[list_index]
    colnames(temp_mat) <- c("Method", "fpr", "tpr", "nullpi", "Nsamp", "alt_type")

    roc_mat <- rbind(roc_mat, temp_mat)
}


for (current_alt_type in alt_type_seq[1:(length(alt_type_seq) - 1)]) {
submat <- dplyr::filter(roc_mat, alt_type == current_alt_type)
p <- ggplot(data = submat, mapping = aes(x = fpr, y = tpr, group = Method, color = Method)) +
    facet_grid(nullpi~Nsamp) +
    geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.5) +
    geom_line(alpha = I(0.7), size = I(0.3)) + xlab("FPR") + ylab("TPR") +
    ggtitle(paste("Pointwise Mean ROC when Alt Type =", current_alt_type)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))
print(p)
}
```

\clearpage

```{r, cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(pROC)

rm(list = ls())
load("betahat_muscle.Rd")
load("betahat_muscle_ash.Rd")
load("betahat_muscle_ruvash.Rd")
load("betahat_muscle_sruvash.Rd")
par_vals <- read.csv("sim_settings.csv")

for (index in 1:length(ash_betahat)) {
    betahat[[index]]$RUVASHOLS <- ash_betahat[[index]][,1]
    betahat[[index]]$RUVASHGLS <- ruvash_betahat[[index]][,1]
    betahat[[index]]$SRUVASHGLS <- sruvash_betahat[[index]][,1]
    betahat[[index]] <- betahat[[index]][, c(1:(ncol(betahat[[index]]) - 4),
                                       ncol(betahat[[index]]) - 2, ncol(betahat[[index]]) - 1,
                                       ncol(betahat[[index]]),
                                       ncol(betahat[[index]]) - 3)]
}


get_mse <- function(beta_mat) {
    which_betatrue <- which(colnames(beta_mat) == "beta_true")
    return(colMeans((beta_mat$beta_true - beta_mat[, -which_betatrue]) ^ 2))
}

mse_mat          <- as.data.frame(t(sapply(betahat, get_mse)))
colnames(mse_mat) <- colnames(betahat[[1]])[1:(length(betahat[[1]]) - 1)]
mse_mat$nullpi   <- par_vals$nullpi
mse_mat$Nsamp    <- par_vals$Nsamp
mse_mat$alt_type <- par_vals$alt_type
alt_type_seq     <- unique(mse_mat$alt_type)

for (alt_type_index in 1:length(alt_type_seq)) {
    current_alt_type <- alt_type_seq[alt_type_index]

    mse_mat_sub <- mse_mat[mse_mat$alt_type == current_alt_type,]
    ymax <- quantile(c(as.matrix(dplyr::select(mse_mat_sub, -c(nullpi, Nsamp, alt_type)))),
             0.999, na.rm = TRUE)
    mse_mat_sub <- dplyr::select(mse_mat_sub, -alt_type)
    long_dat <- melt(mse_mat_sub, id.vars = c("nullpi", "Nsamp"))
    p <- ggplot(data = long_dat, mapping = aes(x = variable, y = value, fill = variable))
    p <- p + facet_grid(nullpi~Nsamp) + ylab("MSE")
    p <- p + geom_boxplot(size = 0.4, outlier.size = 0.4)
    p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))
    p <- p + ggtitle(paste("MSE When Using Muscle Tissue, Alternative =",
                           alt_type_seq[alt_type_index])) +
        ylim(0, ymax)
    print(p)
}

```

\clearpage

# MSE without LEAPP

```{r, echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE}

mse_mat <- dplyr::select(mse_mat, -c(leapp_sparse, leapp_ridge))
for (alt_type_index in 1:length(alt_type_seq)) {
    current_alt_type <- alt_type_seq[alt_type_index]

    mse_mat_sub <- mse_mat[mse_mat$alt_type == current_alt_type,]
    mse_mat_sub <- dplyr::select(mse_mat_sub, -alt_type)
    long_dat <- melt(mse_mat_sub, id.vars = c("nullpi", "Nsamp"))
    p <- ggplot(data = long_dat, mapping = aes(x = variable, y = value, fill = variable))
    p <- p + facet_grid(nullpi~Nsamp) + ylab("MSE")
    p <- p + geom_boxplot(size = 0.4, outlier.size = 0.4)
    p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))
    p <- p + ggtitle(paste("MSE When Using Muscle Tissue, Alternative =",
                           alt_type_seq[alt_type_index]))
    print(p)
}
```


```{r}
sessionInfo()
```
